Stateless Model Checking with Data-Race Preemption Points

### 2015-10-02 pldi-16/evaluation.tex
some touchups and talk about test suite

Although \quicksand~presents Iterative Deepening and data-race PPs as interconnected techniques, they each could theoretically be employed alone in other model checkers.
For example, a single-state-space tool could use data-race candidates during immediately subsequent interleavings, essentially changing the state space on the fly.
Likewise, a message-passing-only tool could employ Iterative Deepening despite data races being absent from its concurrency model.
Hence, though many of our experiments compare \quicksand~to the state-of-the-art as a whole,
we also sought to evaluate each technique individually.
Our evaluation answers the following questions:
\begin{enumerate}
	\item Is combining data-race analysis with stateless model checking productive?
		\begin{enumerate}
			\item Do data-race PPs expose new bugs that couldn't be found with single-state-space model checking?
				% Elaborate later:
				% Among those, how many were missed in a {\em completed} execution of the otherwise ``maximal'' state space?
			\item Do we avoid false positives compared to single-pass?
				% Explain later as:
				% How many data-race candidates were verified as benign
				% Also Include:
				% How many false positives does the free-re-malloc technique suppress?
			% TODO: This one's optional. You can give up on it.
			\item Do we avoid false negatives compared to single-pass?
				% Explain later as: How many data-race bugs required backtracking to be identified as candidates in the first place
				%(don't forget set EXPLORE_BACKWARDS=0 here)
		\end{enumerate}
	\item Is Iterative Deepening better than the single-state-space model?
		\begin{enumerate}
			\item Does Iterative Deepening find bugs
				% in subset state spaces
				faster than the single-state-space approach, even without data races?
			\item Does Iterative Deepening help test data-race PPs?
				% Are there some DRs where the PP alone finds the bug, while the superset SS cannot, and some vice versa?
				% What we want is a *variety*, which proves that iterative deepening is better than either individual strategy.
		\end{enumerate}
\end{enumerate}

>>>>

Although \quicksand~presents Iterative Deepening and data-race PPs as interconnected techniques, they each could theoretically be employed alone in other model checkers.
For example, a single-state-space tool could use data-race candidates during immediately subsequent interleavings, essentially changing the state space on the fly.
Likewise, a message-passing-only tool could employ Iterative Deepening despite data races being absent from its concurrency model.
Hence, though many of our experiments compare \quicksand~to the state-of-the-art as a whole,
we also sought to evaluate each technique individually.
Our evaluation answers the following questions:
\begin{enumerate}
	\item Is combining data-race analysis with stateless model checking productive?
		\begin{enumerate}
			\item Do data-race PPs expose new bugs that couldn't be found with single-state-space model checking?
				% Elaborate later:
				% Among those, how many were missed in a {\em completed} execution of the otherwise ``maximal'' state space?
			\item Do we avoid false positives compared to single-pass?
				% Explain later as:
				% How many data-race candidates were verified as benign
				% Also Include:
				% How many false positives does the free-re-malloc technique suppress?
			% TODO: This one's optional. You can give up on it.
			\item Do we avoid false negatives compared to single-pass?
				% Explain later as: How many data-race bugs required backtracking to be identified as candidates in the first place
				%(don't forget set EXPLORE_BACKWARDS=0 here)
		\end{enumerate}
	\item Is Iterative Deepening better than the single-state-space model?
		\begin{enumerate}
			\item Does Iterative Deepening find bugs
				% in subset state spaces
				faster than the single-state-space approach, even without data races?
			\item Does Iterative Deepening help test data-race PPs?
				% Are there some DRs where the PP alone finds the bug, while the superset SS cannot, and some vice versa?
				% What we want is a *variety*, which proves that iterative deepening is better than either individual strategy.
		\end{enumerate}
\end{enumerate}
\subsection{Test Suite}
% TODO: Maybe say how many lines of code total? How many lines on average per P2/pintos? (careful with pintos; lots of basecode)
Our test suite consists of \numthrlibs~``P2'' student thread libraries, from CMU's 15-410 OS class,
%across the Spring and Fall 2014 and Spring 2015 semesters;
and \numpintoses~``Pintos'' student kernels, from Berkeley's CS162 and U. Chicago's CS230 OS classes.
%
The P2 thread library comprises \texttt{thr\_create()}, \texttt{thr\_exit()}, \texttt{thr\_join()}, mutexes, condvars, semaphores, and r/w locks;
all implemented from scratch in userspace with a UNIX-like system call interface \cite{kspec,thrlib}.
%
The Pintos kernel project
involves implementing priority scheduling, \texttt{sleep()}, and user-space process management (\texttt{wait()} and \texttt{exit()})
using provided bare-bones mutex, context-switch, and virtal memory implementations
\cite{pintos}.

% TODO: List test programs that were run.

### 2015-10-02 pldi-16/intro.tex
some touchups and talk about test suite

We evaluated \quicksand~by testing
% TODO: What's the real number
{\Large \em 31337}
student thread libraries and kernels from the undergraduate OS classes at Carnegie Mellon, Berkeley, and the University of Chicago.
We show that \quicksand~finds more bugs than the conventional stateless model checking approach given the same CPU budget,
and furthermore that adding data-race preemption points quickly exposes bugs missed by even the ``maximal'' state space of the conventional approach.

>>>>

We evaluated \quicksand~by testing \numstudence~student thread libraries and kernels from the undergraduate OS classes at Carnegie Mellon, Berkeley, and the University of Chicago.
We show that \quicksand~finds more bugs than the conventional stateless model checking approach given the same CPU budget,
and furthermore that adding data-race preemption points quickly exposes bugs missed by even the ``maximal'' state space of the conventional approach.

### 2015-10-02 pldi-16/related.tex
some touchups and talk about test suite

\quicksand~is built upon many established concurrency testing techniques, dating back of course to Verisoft, the original C model checker \cite{verisoft}.
We use \landslide~as our model checker \cite{landslide},
% TODO: is DPOR abbreviated earlier in the paper
which itself implements DPOR \cite{dpor},
state space estimation \cite{estimation},
and data race detection \cite{eraser}. Related model checkers can largely be classified based on their treatment of shared-memory thread communication.

{\bf Synchronization events only.} CHESS \cite{chess} and \textsc{dBug} \cite{dbug-ssv} instrument the thread library API, and can preempt programs only during calls to this API.
Hence they will miss any crashes resulting from interleaving threads at instruction granularity during a data race. CHESS provides a data-race analysis to report any such violations of its concurrency model to the user, but does not incorporate data race candidates as PPs in future tests.

{\bf Message-passing.} Other stateless model checkers, such as SAMC \cite{samc}, \textsc{MaceMC} \cite{macemc}, \textsc{MoDist}, and ETA \cite{parallel-dpor}, require thread communication to be restricted to a message-passing API to more effectively test distributed systems.
This eliminates the need for data race analysis, but restricts the class of programs that can be tested.
Nevertheless, Iterative Deepening is still applicable to these tools.

@@

{\bf Other techniques.} Various improvements to DPOR have been proposed, such as Dynamic Interface Reduction \cite{demeter}, Maximal Causality Reduction \cite{mcr}, and DPOR for TSO/PSO \cite{tsopso}.
These are all orthogonal to our technique.
\textsc{Parrot} \cite{parrot} combines model checking with a partially-determinizing runtime environment, but still, fewer than half the non-trivial state spaces in their evaluation could be completed, providing a strong argument for \quicksand.
Finally, Iterative Context Bounding \cite{chess} is most similar to Iterative Deepening, as both approaches provide a concrete partial verification on some subset of interleavings when full completion is intractable.
However, ICB still limits itself to a fixed set of PPs, and so far no algorithm has been proposed to dynamically add data race PPs during a test with ICB.

>>>>

\quicksand~is built upon many established concurrency testing techniques, dating back
% of course
to Verisoft, the original C model checker \cite{verisoft}.
Our model checker \landslide~\cite{landslide}
% TODO: is DPOR abbreviated earlier in the paper
itself implements DPOR \cite{dpor},
state space estimation \cite{estimation},
and data-race detection \cite{eraser}.
We classify related model checkers largely based on their treatment of shared-memory thread communication.

{\bf Synchronization events only.} CHESS \cite{chess} and \textsc{dBug} \cite{dbug-ssv} instrument the thread library API, and can preempt programs only during calls to this API.
Hence they will miss any crashes that require interleaving threads at instruction granularity during a data race. CHESS provides a data-race analysis to report any such violations of its concurrency model to the user, but does not incorporate data-race candidates as PPs in future tests.

{\bf Message-passing.} Other stateless model checkers, such as SAMC \cite{samc}, \textsc{MaceMC} \cite{macemc}, \textsc{MoDist}, and ETA \cite{parallel-dpor}, require thread communication to be restricted to a message-passing API to more effectively test distributed systems.
This eliminates the need for data-race analysis, but restricts the class of programs that can be tested.
Nevertheless, Iterative Deepening is still applicable to these tools.

@@

{\bf Other techniques.} Various improvements to DPOR have been proposed, such as Dynamic Interface Reduction \cite{demeter}, Maximal Causality Reduction \cite{mcr}, and DPOR for TSO/PSO \cite{tsopso}.
These are all orthogonal to our technique.
\textsc{Parrot} \cite{parrot} combines model checking with a partially-determinizing runtime environment, but still, fewer than half the non-trivial state spaces in their evaluation could be completed, providing a strong argument for \quicksand.
Finally, Iterative Context Bounding \cite{chess} is most similar to Iterative Deepening, as both approaches provide a concrete partial verification on some subset of interleavings when full completion is intractable.
However, ICB still limits itself to a fixed set of PPs, and so far no algorithm has been proposed to dynamically add data-race PPs during a test with ICB.
